Data Management for LSST Special Programs

### 2024-03-01 ap_procbounds.tex
Apply suggestions from code review
Co-authored-by: Leanne Guy <5301176+leannep@users.noreply.github.com>

\begin{itemize}
\item \bf{Short or long exposures.} Very short ($<$2 sec) exposures could be difficult to 
process due to an incompletely-formed PSF. Short (or long, $>$150 sec) exposures 
could be difficult to calibrate due to having too few (or too few unsaturated) stars.
\item \bf{Bright backgrounds (twilight).} It is currently unclear whether images with very 
bright sky backgrounds (twilight images) can be processed with the LSST Science Pipelines, 
or whether user generated pipelines will be needed.
\item \bf{Non-sidereal tracking.} The full reduction and calibration of images obtained 
with non-sidereal tracking, in which the stars are streaked, is currently beyond the scope 
of the LSST Science Pipelines, and will require a user generated pipeline.
\end{itemize}

>>>>

\begin{itemize}
\item \textbf{Short or long exposures.} Very short ($<$2 sec) exposures could be difficult to 
process due to an incompletely-formed PSF. Short (or long, $>$150 sec) exposures 
could be difficult to calibrate due to having too few (or too few unsaturated) stars.
\item \textbf{Bright backgrounds (twilight).} It is currently unclear whether images with very 
bright sky backgrounds (twilight images) can be processed with the LSST Science Pipelines, 
or whether user generated pipelines will be needed.
\item \textbf{Non-sidereal tracking.} The full reduction and calibration of images obtained 
with non-sidereal tracking, in which the stars are streaked, is currently beyond the scope 
of the LSST Science Pipelines, and will require a user generated pipeline.
\end{itemize}

### 2024-03-01 sec_intro.tex
Apply suggestions from code review
Co-authored-by: Leanne Guy <5301176+leannep@users.noreply.github.com>

\subsection{Main Survey and Wide-Fast-Deep (WFD)}

@@

This is a term used only in this document to refer to the image processing 
described in the Data Products Definitions Document
(LSE-163) that are designed for, and will be applied to, the Main Survey's observations.

@@

This is a Rubin Data Management term used to refer to sky regions within
or beyond the Main Survey footprint which have a distinct observing strategy from
the Main Survey's plans for that region (e.g., different visit type, cadence, filter distribution).

>>>>

\subsection{Main Survey including Wide-Fast-Deep (WFD)}

@@

This is a term used only in this document to refer to the image processing 
described in the Data Management Science Pipelines Design \citeds{LDM-151} that produces the data products described in the Data Products Definitions Document
(\citeds{LSE-163}) that are designed for, and will be applied to, the Main Survey's observations.

@@

This is a Rubin Data Management term used to refer to sky regions within
or beyond the Main Survey footprint that have a distinct observing strategy from
the Main Survey's plans for that region (e.g., different visit type, cadence, filter distribution).

### 2024-03-01 sec_proc.tex
Apply suggestions from code review
Co-authored-by: Leanne Guy <5301176+leannep@users.noreply.github.com>

\item \textbf{Short exposures --}
Special Programs that use short exposures would be limited to the
minimum exposure time is 1 second (stretch goal: 0.1 seconds\footnote{A short-exposure survey 
of the bright stars of M67, described in Chapter 10.4 of the 
Observing Strategy White Paper \citep{2017arXiv170804058L}, suggests using the stretch goal of 
0.1 second exposures or, if that is not possible, \textit{"custom pixel masks to accurately perform 
photometry on stars as much as 6 magnitudes brighter than the saturation level"}. 
This may require User-Generated Processing.}).
There is a potential hardware boundary that limits the readout rate to 1 
every 15 seconds, which would affect the image acquisition rate and 
increase the overheads on short exposures.

\item \textbf{Repeated pointing --}
Special Programs which require the \emph{exact same} field pointing and 
rotation for \emph{every exposure} (to sub-arcsecond levels) might run 
into hardware boundaries on pointing and tracking.

@@

\item \textbf{Very short exposures --}
Special Programs which use very short ($<$2 sec) exposures 
could be difficult to reduce due to an incompletely-formed PSF 
(Section~\ref{ssec:procbounds_expt}).
The Data Management System is required to be able to process exposure 
times as low as 1 second (Section~\ref{ssec:req_proc}), 
but it is known that such short exposures might have degraded image quality.

@@

\item \textbf{Twilight images --}
For Special Programs which obtain images with very bright sky backgrounds
(twilight images), it is currently unclear whether they can be processed
with the LSST Science Pipelines; 
User-Generated Processing might be needed (e.g., \citealt{2022AJ....164..168S}).

@@

Decisions about when to apply regular processing to Special Programs data,
or when to include it in the data products for the Main Survey
(e.g., if it improves the all-sky coadds), 
are ultimately left to the discretion of the Rubin Observatory's 
Data Management and System Performance teams.

@@

Options for RDM to process TOO observations, especially during the first year of Operations 
when the template coverage will be low, are discussed in more detail in 
\citeds{rtn-008}.

>>>>

\item \textbf{Short exposures --}
Special Programs that use short exposures would be limited to the
minimum exposure time of 1 second (stretch goal: 0.1 seconds\footnote{A short-exposure survey 
of the bright stars of M67, described in Chapter 10.4 of the 
Observing Strategy White Paper \citep{2017arXiv170804058L}, suggests using the stretch goal of 
0.1 second exposures or, if that is not possible, \textit{"custom pixel masks to accurately perform 
photometry on stars as much as 6 magnitudes brighter than the saturation level"}. 
This may require User-Generated Processing.}).
There is a potential hardware boundary that limits the readout rate to 1 
every 15 seconds, which would affect the image acquisition rate and 
increase the overheads on short exposures.

\item \textbf{Repeated pointing --}
Special Programs that require the \emph{exact same} field pointing and 
rotation for \emph{every exposure} (to sub-arcsecond levels) might run 
into hardware boundaries on pointing and tracking.

@@

\item \textbf{Very short exposures --}
Special Programs that use very short ($<$2 sec) exposures 
could be difficult to reduce due to an incompletely-formed PSF 
(Section~\ref{ssec:procbounds_expt}).
The Data Management System is required to be able to process exposure 
times as low as 1 second (Section~\ref{ssec:req_proc}), 
but it is known that such short exposures might have degraded image quality.

@@

\item \textbf{Twilight images --}
For Special Programs that obtain images with very bright sky backgrounds
(twilight images), it is currently unclear whether they can be processed
with the LSST Science Pipelines; 
User-Generated Processing might be needed (e.g., \citealt{2022AJ....164..168S}).

@@

Decisions about when to apply regular processing to Special Programs data,
or when to include it in the data products for the Main Survey
(e.g., if it improves the all-sky coadds), 
are ultimately left to the discretion of the Rubin Observatory's 
Data Management and System Performance departments in Rubin Operations.

@@

Options for Data Management to process TOO observations, especially during the first year of Operations 
when the template coverage will be low, are discussed in more detail in 
\citeds{rtn-008}.

### 2024-03-01 sec_sci.tex
Apply suggestions from code review
Co-authored-by: Leanne Guy <5301176+leannep@users.noreply.github.com>

\textbf{Necessary -- } When the primary science goal for a Special Program 
cannot be met by including the data in regular processing (e.g., Prompt processing),
or where doing so would compromise the Main Survey data products (e.g., introduce non-uniformity).

@@

Users will need to be able to query for data that restricts by sky region and 
observign mode label. 
This can be accomplished by including those labels in all image and catalog 
metadata as described in Section~\ref{ssec:sci_labels}, as the TAP service 
and butler already provide the mechanism for user-specified queries.

>>>>

\textbf{Necessary -- } When the primary science goal for a Special Program 
cannot be met by including the data in regular processing (e.g., Prompt processing),
or where doing so would compromise the Main Survey data products (e.g., introduce additional non-uniformity beyond what would be expected based on the survey strategy).

@@

Users will need to be able to query for data that restricts by sky region and 
observing mode label. 
This can be accomplished by including those labels in all image and catalog 
metadata as described in Section~\ref{ssec:sci_labels}, as the TAP service 
and butler already provide the mechanism for user-specified queries.
